---
output: pdf_document
---
#Automatic vs Manual MPG Regression Models

##Executive Summary

I am working for Motor Trend, a magazine about the auto industry. This paper will examine the relationship between a set of variables and miles per gallon (MPG) found in the mtcars data set provided by R. In particular I am looking at the questions 

-Is automatic or manual transmission better for MPG
-Quantify this difference

##Preliminary Investigation and Exploratory Analysis

First I'll load the data into R. Then I'll print the variables in the data set so we can see what we're working with. 

```{r}
data(mtcars)
names(mtcars)
```

The "am" column is the transimission (0 = automatic, 1 = manual). I want to use that to predict mpg. I'll turn "am" into a factor (and label it appropriately) and make a boxplot to get a sense of the data

```{r}
mtcars$am <- as.factor(mtcars$am)
levels(mtcars$am) <- c("Automatic", "Manual")
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission Type", ylab = "MPG", main = "MPG Distribution by Transmission")
```

That's interesting. So it seems at a glance that manual transimissions tend to have better mileage than automatic. 

##Basic Linear Model

Since we're interested in transmissions, I'll fit a linear model based purely on the transimission. 
```{r}
fit <- lm(mpg ~ am, data = mtcars)
summary(fit)
```

R has automatically picked "Automatic" transimission cars as the baseline, so the Intercept of 17.147 is actually the average mpg for automatic cars. The coefficient for the "Manual" transmission is interepreted as "how many more miles per gallon on average do you get by switching from an Automatic to a Manual" which is 7.245. This is consistent with our earlier findindgs, it seems that "Manual" transmission cars on average get 7.245 MORE miles to the gallon than automatic cars do. However, note the Adjusted R-Squared is only 0.3385, which is quite low. This means this model does not fit the data terribly well. 

##Finding Improved Linear Model

Now let's find a better fit for our model. I'm going to update the original fit model with each variable in the mtcars set one at a time, building more complex models. I will then examine the Pr(>F) value and if it is statistically significant (less than 0.05) this means that adding the additional variable provides a better "significantly" better fit to the data.

```{r}
fit2 <- update(fit, mpg ~am + wt)
fit3 <- update(fit, mpg ~am + wt + qsec)
fit4 <- update(fit, mpg ~ am + wt + qsec + cyl)
anova(fit, fit2, fit3, fit4)
```

It is clear that the fit3 model, mpg being predicted by am, wt, and qsec gives us the strongest evidence to reject the null hypothesis. I have not included the other fits for the sake of brevity, but they do not fit the data significantly better.  

```{r}
summary(fit3)
```

This is much better. Now we have an adjusted R squared of 0.8336, meaning this model explains 88.36% of the variance making for a much better fit. Also the coefficients all have small p-values (< 0.05), so they are statistically significant (although the Intercept is not). Now I'll examine some interactions to see if they should be included. 

```{r}
bestfit <- update(fit, mpg ~ am + wt + qsec + am*wt)
anova(fit3, bestfit)
```

After trial and error (not detailed here for sake of space) I have discovered that the interaction am*wt should be included to create my bestfit model. The F test is clearly small, meaning the change is significant. 

```{r}
summary(bestfit)
```

Now we can see we have an adjusted R squared of 0.8804 and small p-values for the coefficients, so this model explains 88.04% of the variance in our data providing a good fit with statistically significant coefficients. 


```{r}
par(mfrow = c(2,2))
plot(bestfit)
```

There does not seem to be a pattern in the residual plot, meaning we have avoided heteroskedasticity (non-constant variance). Thus bestfit seems to be a good overall model for explaining mileage

##Conclusions

So what does this mean? Well if we look at the difference between an automatic and manual car, a manual transmission car gets 14.079 more miles to the gallon than an Automatic - 4.141 * the weight. 

```{r}
14.079 - 4.141 * mean(mtcars$wt)
```

So we can see on average a Manual car only gets 0.76 more miles to the gallon than an Automatic (all other things being equal). This is not terribly much. But again this is for averages, if a car is light weight it probably makes sense to get a manual transmission to get better mileage (under this model)

##Further Investigation and Notes

It would be interesting to see how this compares model performs on a training set. Mtcars only contains 32 observations, making this a very small sample. This data also comes from the 1974 issue of Motor Trend, which is not quite up to date with modern models. 